Runtime Feedback in a Meta-Tracing JIT for Efficient Dynamic Languages

### 2011-04-04 talk/icooolps2011/paper.tex
swap these two paragraphs; I think it is more readable this way
another small tweak
try to take the best of both versions

% XXX: paragraph rephrased by anto; feel free to pick the one you like best

Conceptually, it is possible to achieve significant speed-ups by feeding into
the compiler some information that is observed at runtime: in particular, if
there are values which vary very slowly, it is possible to compile multiple
specialized versions of the same code, one for each actual value.  To exploit
the runtime feedback, the implementation code and data structures need to be
structured so that many such slow-varying values are at hand. The hints that
we present allow exactly to implement such feedback and exploitation in a
meta-tracing context.

% Conceptually the significant speed-ups that can be achieved with
% dynamic compilation depend on feeding into compilation and exploiting
% values observed at runtime that are slow-varying in practice. To exploit the
% runtime feedback, the implementation code and data structures need to be
% structured so that many such values are at hand. The hints that we present allow
% exactly to implement such feedback and exploitation in a meta-tracing context.

@@

While the operations in a trace are those of the interpreter, the loops that are
traced by the tracer are the loops in the
user program. This means that the tracer stops tracing after one iteration of
the loop in the user function that is being considered. At this point, it can
have traced many iterations of the interpreter main loop.

@@

\subsection{Running Example}
\label{sub:running}

@@

When using this object model in
an interpreter, a huge amount of time will be spent doing lookups in these
dictionaries.
Let's assume we trace through code that sums three attributes, such as:

@@

Promotion is essentially a tool for trace specialization. In some places in the
interpreter it would be very useful if a variable were constant, even though it
could have different values in practice. In such a place, promotion is used. The
typical reason to do that is if there is
a lot of computation depending on the value of that variable.

@@

Why can't we simply write an analysis to find out that the \texttt{x} fields of the
\texttt{A} instances is immutable and deduce that \texttt{compute} is a pure function,
since it only reads the \texttt{x} field and does not have side effects? This might
be possible in this particular case, but in practice the functions that are
annotated with the \texttt{purefunction} decorator are usually more complex.
The easiest example for this is that of a function that uses memoization to
cache its results. If this function is analyzed, it looks like the function has
side effects, because it changes the memoizing dictionary. However, because this side
effect is not externally visible, the function from the outside is pure. This is
a property that is not easily detectable by analysis. Therefore, the purity
of this function needs to be annotated.

>>>>

Conceptually, the significant speed-ups that can be achieved with
dynamic compilation depend on feeding into compilation and exploiting
values observed at runtime. In particular, if
there are values which vary very slowly, it is possible to compile multiple
specialized versions of the same code, one for each actual value.  To exploit
the runtime feedback, the implementation code and data structures need to be
structured so that many such slow-varying values are at hand. The hints that
we present allow exactly to implement such feedback and exploitation in a
meta-tracing context.

@@

While the operations in a trace are those of the interpreter, the loops that are
traced by the tracer are the loops in the
user program. This means that the tracer stops tracing after one iteration of
the loop in the user function that is being considered. At this point, it probably
traced many iterations of the interpreter main loop.

@@

\subsection{Running Example}
\label{sub:running}
\anto{this example is not referenced until section \ref{sec:fastobjmodel}: I
  would put it just before it}

@@

When using this object model in
an interpreter, a huge amount of time will be spent doing lookups in these
dictionaries.
Let's assume we trace through code that sums three attributes, such as:
\anto{I still think it's a bit weird to call them ``methods'' and then use
  them as attributes in the example}

@@

Promotion is essentially a tool for trace specialization. There are places in
the interpreter where knowing that a value is constant opens a lot of
optimization opportunities, even though it
could have different values in practice. In such a place, promotion is used. The
typical reason to do that is if there is
a lot of computation depending on the value of that variable.

@@

Why can't we simply write an analysis to find out that the \texttt{x} fields of the
\texttt{A} instances is immutable and deduce that \texttt{compute} is a pure function,
since it only reads the \texttt{x} field and does not have side effects? This might
be possible in this particular case, but in practice the functions that are
annotated with the \texttt{purefunction} decorator are usually more complex.
The easiest example for this is that of a function that uses memoization to
cache its results. If this function is analyzed, it looks like the function has
side effects, because it changes the memoizing dictionary. However, because this side
effect is not externally visible, the function from the outside is pure. This is
a property that is not easily detectable by analysis. Therefore, the purity
of this function needs to be annotated manually.
